### 2016-10-28 libcxx-slides.tex
libcxx-slides

\documentclass{beamer}
\usepackage{hyperref}
\begin{document}

\def \GVN {GVN}

\title{libcxx}
\author{Sebastian Pop and Aditya Kumar}
\institute{SARC: Samsung Austin R\&D Center}
\date{\today}

\frame{\titlepage}

\frame{\frametitle{libcxx-performance improvements}
  \begin{itemize}
  \item PR21192 - Reading from stdin is 1-2 orders of magnitude slower than using libstdc++
    - implementation of sync\_with\_stdio doesn't disable synchronization \url{http://stackoverflow.com/q/33837349}

  \item PR19708 - std::find is significantly slower than libstdc++
    - using Duff's device
    - unrolling the loop
    - enabling vectorization \url{http://stackoverflow.com/q/2741859}

  \item PR20837 - libc++'s std::sort is O(N^2) in the worst case (instead of O(N*ln(N))).
    - switching to heapsort for large size

  \item PR26886 - libc++'s std::stable\_sort also has a worst-case complexity issue

  \item PR15456 - A faster implementation of std::function is possible

  \item PR16747 and PR21275 - Our unordered\_multimap insert is much slower than libstdc++'s

  \item Removing Heap usage

  \item Missing features (and performance issues) in <regex>

  \item Non-conforming allocators in the test suite

  \item Identifying missing attributes in functions

  \item Debug mode for the library
  \end{itemize}
}

\end{document}