### 2018-02-23 cgo/paper.tex
cc comments

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,10pt,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,10pt,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall,10pt]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
%\acmJournal{PACMPL}
%\acmVolume{1}
%\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
%\acmArticle{1}
%acmYear{2017}
%\acmMonth{1}
%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%\setlength{\abovecaptionskip}{5pt plus 1pt minus 1pt} % Chosen fairly arbitrarily
%\setlength{\belowcaptionskip}{5pt plus 1pt minus 1pt} % Chosen fairly arbitrarily

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{comment}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{graphviz}
\usepackage{auto-pst-pdf}
\usepackage{etoolbox}
\usepackage{flushend}
\usepackage{needspace}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  %columns=fixed,
  %fontadjust=true,
  basewidth=0.4em,
  language=C,
  mathescape
}
\usepackage[final]{pdfpages}

\renewcommand\lstlistingname{Algorithm}

\begin{document}

%% Title information
\title[SSA GCM]{A fast algorithm for global code motion of congruent instructions}         %% [Short Title] is optional;

\def \GCC {GCC}
\def \LLVM {LLVM}
\def \CFG {CFG}
\def \SSA {SSA}
\def \MemorySSA {MemorySSA}
\def \PRE {PRE}
\def \GVN {GVN}
\def \SPEC {SPEC CPU 2006}
\def \gcm {global-code-motion}
\def \GCM {GCM}
\def \xlinux {x86\_64-linux}
\def \ooo {out-of-order}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Aditya Kumar}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Senior Compiler Engineer}
  %\department{Department1}              %% \department is recommended
  \institution{Samsung Austin R\&D Center}            %% \institution is required
  %\streetaddress{Street1 Address1}
  \city{Austin}
  \state{Texas}
  %\postcode{Post-Code1}
  \country{USA}
}
\email{aditya.k7@samsung.com}

\author{Sebastian Pop}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Senior Staff Compiler Engineer}
  %\department{Department1}              %% \department is recommended
  \institution{Samsung Austin R\&D Center}            %% \institution is required
  %\streetaddress{Street1 Address1}
  \city{Austin}
  \state{Texas}
  %\postcode{Post-Code1}
  \country{USA}
}
\email{s.pop@samsung.com}

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
We present a fast \gcm{} (\GCM{}) compiler optimization which schedules
congruent instructions across the program. We introduce data structures to help
exploit flow of global values in sparse representation.  Not only \GCM{}
improves performance it saves code size as well, it exposes more instruction
level parallelism in the basic-block to which instructions are moved, and it
enables other compiler optimizations like loop invariant motion to remove
redundancies. The cost model to drive the code motion is based on liveness
analysis on \SSA{} representation such that the (virtual) register pressure does
not increase resulting in 2\% fewer spills on the \SPEC{} benchmark
suite. \GCM{} is an optimistic algorithm in the sense that it considers all
congruent instructions in a function as potential candidates. We also formalize
why register pressure reduces as a result of \gcm{}, and how \gcm{} increases
instruction level parallelism thereby enabling more \ooo{} execution on modern
architectures.

We have implemented the pass in \LLVM{}. The experimental results show
performance gains of 8\% in 462.libquantum and 5.3\% in 471.omnetpp from the
\SPEC{} benchmark suite. It also reduces compilation time by reducing number of
instructions to be processed by later compilation stages. \GCM{} enables more
inlining and exposes more loop invariant code motion opportunities in the
majority of benchmarks.
\end{abstract}

%% Keywords
%% comma separated list
\keywords{Optimizing Compilers, \GCC{}, \LLVM{}, Code Generation, Global Scheduling}  %% \keywords is optional

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
\label{sec:intro}
Compiler techniques to remove redundant computations are composed of an analysis
phase that detects identical computations in the program and a transformation
phase that reduces the number of run-time computations.  Classical scalar
optimizations like Common Subexpression Elimination (CSE) \cite{dragonbook} work
very well on single basic-blocks (local level) but when it comes to detect
redundancies across basic-blocks (global level) these techniques fall short:
more complex passes like Global Common Subexpression Elimination (GCSE) and
Partial Redundancy Elimination (\PRE{}) have been designed to handle these cases
based on data-flow analysis \cite{morel1979global}.  At first these techniques
were described in the classical data-flow analysis framework, but later the use
of Static Single Assignment (\SSA{}) representation lowered their cost in terms
of compilation time \cite{briggs1994effective,chow1997new,kennedy1999partial}
and brought these techniques in the main stream: \SSA{} based \PRE{} has been
implemented in many optimizing compilers like \GCC{} and \LLVM{}.

This paper describes a fast algorithm for \gcm{} (\GCM{}) of congruent
instructions \cite{briggs1997}, a technique that uses the information computed
for \PRE{} to detect identical computations but has a transformation phase whose
goal differs from \PRE{}: it removes identical computations from different
branches of execution.  These identical computations in different branches of
execution are not redundant computations at run-time and the number of run-time
computations is not reduced. It is not a redundancy elimination pass, and thus
it has different cost function and heuristics than \PRE{} or CSE. It is similar
to global code scheduling \cite{dragonbook,click1995global} in the sense that it
will only move dynamic computations. Code hoisting, for example, can reduce the
critical path length of execution in \ooo{} machines. As more instructions are
available at the hoisting point, the hardware has more independent instructions
to reorder. The following reduced example illustrates how hoisting can improve
performance by exposing more ILP.

\begin{lstlisting}
float fun(float d, float min, float max, float a) {
  float tmin, tmax;
  float inv = 1.0f / d;
  if (inv >= 0) {
    tmin = (min - a) * inv;
    tmax = (max - a) * inv;
  } else {
    tmin = (max - a) * inv;
    tmax = (min - a) * inv;
  }
  return tmax + tmin;
}
\end{lstlisting}

In this program the computations of $tmax$ and $tmin$ are identical to the
computations of $tmin$ and $tmax$ of sibling branch respectively. Both $tmax$ and $tmin$
depend on $inv$ which depends on a division operation which is generally more
expensive than the addition, subtraction, and multiplication operations. The
total latency of computation across each branch is: $C_{div} + 2(C_{sub} + C_{mul})$.
For an \ooo{} processor with two add units and two multiply
units, the total latency of computation is $C_{div} + C_{sub} + C_{mul}$.

Now if the computation of $tmax$ and $tmin$ are hoisted outside the
conditionals, the C code version would look like:
\begin{lstlisting}
float fun(float d, float min, float max, float a) {
  float inv = 1.0f / d;
  float x = (min - a) * inv;
  float y = (max - a) * inv;
  float tmin = x, tmax = y;
  if (inv < 0) { tmin = y; tmax = x; }
  return tmax + tmin;
}
\end{lstlisting}

In this code the division operation can be executed in parallel with the two
subtractions because there are no dependencies among them. So the total number
of cycles will be $max(C_{div}, C_{sub}) + C_{mul} = C_{div} + C_{mul}$; since
$C_{div}$ is usually much greater than $C_{sub}$ \cite{x86,aarch64}.  Although
it decreases the instruction level parallelism (ILP) of basic blocks from where
the instructions were removed, those basic blocks cannot execute for more cycles
than the destination basic block. We have seen the performance of a proprietary
benchmark (on an \ooo{} processor) improve by $15\%$.  In fact, there are
several advantages of \GCM{}:
\begin{itemize}[leftmargin=*,topsep=0pt]
\item it helps reduce the code size of the program by replacing multiple
  instructions with one thereby making the function cheaper to inline because
  inliner heuristics depend on instruction count;
\item it exposes more ILP to the later compiler passes. By hoisting identical
  computations to be executed earlier, instruction schedulers can move heavy
  computations earlier in order to avoid pipeline bubbles.
\item it reduces branch misprediction penalty on \ooo{} processors with
  speculative execution of branches: by hoisting or sinking expressions out of
  branches, it can effectively reduce the amount of code to be speculatively
  executed;
\item it reduces interference or register pressure with an appropriate
  cost-model: we have used a cost model in our implementation which only hoists
  or sinks when the register pressure is reduced;
\item it reduces the total number of instructions to be executed for SIMD
  architectures which execute all code in branches based on masking or
  predication \cite{simd};
\item it may improve loop vectorization by reducing a loop with control flow to
  a loop with a single basic-block, should all the instructions in a conditional get
  hoisted or sunk;
\item it enables more loop invariant code motion (LICM): as LICM passes, in
  general, cannot effectively reason about instructions within conditional
  branches in the context of loops, code-motion is needed to move instructions
  out of conditional expressions and expose them to LICM.
\end{itemize}

The algorithm for \GCM{} uses several common representations of the
program that we shortly describe below:
\begin{itemize}[leftmargin=*,topsep=0pt]
\item Dominance (DOM) and Post-Dominance (PDOM) relations \cite{dragonbook} on a
  Control Flow Graph (\CFG{});
\item DJ-Graph \cite{Sreedhar1996} is a data structure that augments the
  dominator tree with join-edges to keep track of data-flow in a program. We use
  DJ-Graph to compute liveness of variables as illustrated in \cite{das2012};
\item Static Single Assignment (\SSA{}) \cite{cytron};
\item Global Value Numbering (\GVN{}) \cite{rosen1988global,click1995global}: to
  identify similar computations compilers use \GVN{}.  Each expression is given
  a unique number and the expressions that the compiler can prove to be
  identical are given the same number;
\item Rank of an expression \cite{rosen1988global}: The expressions are ordered
  according to the amount of dependency they have on other expressions.  For
  example in the expression $a = b+c$, $a$ will have higher rank than rank of
  $b$, and $c$.
\item \MemorySSA{} \cite{novillo2007memory}: it is a factored use-def chain of
  memory operations that the compiler is unable to prove
  independent. \MemorySSA{} accelerates the access to the alias analysis
  information.
\end{itemize}

Dominator tree, Global Value Numbers and Memory SSA are already available in
\LLVM{} but there is no facility to infer liveness of virtual registers. So we
implemented DJ-Graph data structure that allowed us to calculate MergeSets which
was used in the liveness analysis during \GCM{}.

The \GCM{} pass can be broadly divided into the following:
\begin{itemize}[leftmargin=*,topsep=0pt]
\item factor the reachability of expressions (Section~\ref{sec:chi}).
\item find candidates (congruent instructions) suitable for \gcm{}
  (Section~\ref{subsec:optimistic}),
\item compute a point in the program where it is both legal
  (Section~\ref{subsec:legality}) and profitable
  (Section~\ref{subsec:cost-models}) to move the code,
\item move the code to hoist point or sink point
  (Section~\ref{subsec:optimistic}),
\item update data structures to continue iterative \gcm{} (Section~\ref{subsec:optimistic}).
\end{itemize}

There has been a lot of work both in industry and academia related to code
hoisting/sinking, and in general global scheduling \cite{click1995global}. Some
relate code-hoisting to code-size optimization \cite{rosen1988global} and many
\cite{barany2013, shobaki2013} use global scheduling to improve
performance. Most of the recent work on global scheduling are done using integer
linear programming (ILP) which results in prohibitively high compile time which
is not suitable for industrial compilers. To the best of our knowledge we have
not found any reference which explored \gcm{} with a cost-model to reduce
register-pressure. The main contributions of this paper are:
\begin{itemize}[leftmargin=*,topsep=0pt]
\item an optimistic fast algorithm to schedule congruent instructions;
\item introducing an analogue of $\phi$ nodes which allows faster tracking of
  anticipable instructions;
\item a cost model to reduce register spills;
\item experimental evaluation of our implementation in \LLVM{}.
\end{itemize}

\section{Related Work}
There are a lot of bug reports in \GCC{} and \LLVM{} bugzillas
\cite{GCCCodeHoistingBugs, LLVMCodeHoistingBugs}, showing the interest in having
a more powerful code hoist transform.  The current \LLVM{} implementation of
code hoisting in SimplifyCFG.cpp is very limited to hoisting from identical
basic-blocks: the instructions of two sibling basic-blocks are read at the same
time, and all the instructions of the blocks are hoisted to the common parent
block as long as the compiler is able to prove that the instructions are
equivalent.  This implementation does not allow for an easy extension: first in
terms of compilation time overhead the implementation is quadratic in number of
instructions to bisimulate and second, the equivalence of instructions is
computed by comparing the operands which is neither general nor scalable.

\citet{click1995global} describes aggressive \gcm{} in sparse representation. It
first schedules all the instructions as early as possible. This results in very
long live ranges which is mitigated by again scheduling all the instructions as
late as possible. They report a speedup of as high as $23\%$ but there is no
cost model to limit amount of scheduling and could often introduce redundant
computations. It is also not mentioned how safety is ensured while hoisting an
instruction when it is not anticipable. The legality checks are only based on
barriers (Section~\ref{subsec:barriers}) and availability of used operands. Even
introducing a simple integer addition can also result in undefined behavior
\cite{undef}.  Except from this paper we found no reference of global scheduling
in sparse representation.

\citet{dhamdhere1988fast} and \citet{steven1997advanced} mention
code hoisting in a data-flow framework. A list of Very Busy Expressions (VBE)
computed which are hoisted in a basic-block where the expression is
anticipable (all the operands are available.) This algorithm would hoist as far
as possible without regarding the impact on register pressure and as such a cost
model will be required.  Also the description of VBE is based on the classic
data-flow model and an adaptation to a sparse \SSA{} representation is required.

\citet{rosen1988global} also briefly discuss hoisting computations with
identical value numbers from immediate successors. Their algorithm iterates on
computations of same rank and move the code with identical computations from the
sibling branch to a common dominator if they are very busy
\cite{steven1997advanced}. However, the cost-model to mitigate register pressure
is missing, also, there is no mention of sinking congruent instructions.

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions
(VBEs) at basic-block B, i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
VBEs to a predecessor.  It uses ANTIC\_IN[B] to know what expressions will be
computed on every path from the basic block B to exit, and can be computed in B.
It uses AVAIL\_OUT[B] to subtract out those values already being computed.  The
cost function is: for each hoist candidate, if all successors of B are dominated
by B, then we know insertion into B will eliminate all the remaining
computations.  It then checks to see if at least one successor of B has the
value available.  This avoids hoisting it way up the chain to ANTIC\_IN[B].  It
also checks to ensure that B has multiple successors, since hoisting in a
straight line is pointless.  The algorithm continues top down the dominator
tree, working in tandem with \PRE{} until no more hoisting is possible.  One
advantage of \GCC{} implementation is that it works in sync with the
\GVN{}-\PRE{} such that when new hoisting opportunities are created by
\GVN{}-\PRE{}, code-hoisting will hoist them.

\citet{barany2013} presented a global scheduler with ILP formulation with
a goal to minimize register pressure. The results they got were not very
promising. It may be because they only used the scheduler for smaller functions
($<1000$ instructions); also, they compiled the benchmarks for ARM-Cortex which
is more resilient to register pressure because it has more registers compared to
X86, for example.

\citet{shobaki2013} also presented a combinatorial global scheduler with
reasonable performance improvements. It is possible that both Barany and
Shobaki's implementation will have similar results when compiled for same target
architecture. Also, both suffer from the same problem, although Shobaki not so
much, of large compile times which would not suit in industrial compilers like
\GCC{} and \LLVM{}. With the algorithm described in the next section, we got a
reduction in register spills, improved inliner heuristics, improved compile
time, and show performance improvements on \SPEC{} benchmarks.

% TODO: References
% Sink instructions
% Ranking of instructions from Zadeck
% Liveness analysis from Ramakrishna
% Experimental results: Register spills, compile time, run time.
\section{Global Code Motion of congruent computations}
\label{sec:gcm}
The \GCM{} uses the following data structures ($\chi$ and $\Phi$ nodes), that
are introduced below. They allow us to factor the reachability of values in an
efficient way. Together with liveness analysis using the DJ-Graph \cite{das2012}
and Memory SSA, the entire implementation of \gcm{} only uses sparse
representations for efficient query and update.

\subsection{Factoring reachability of expressions}
\label{sec:chi}
Our algorithm factors out the reachability of values such that multiple queries
to find reachability of values are fast. This is based on finding the dominance
frontiers (DF) and the post-dominance frontiers (PDF) of the CFG. These points,
being structural properties of a graph, do not change during the code-motion. A
post-dominance frontier is the basic block where anticipability (ANTIC) of an
instruction (I) is checked because it is the basic block on which the basic
block containing I is control dependent.

So we introduce a data structure ($\chi$ nodes) to keep track of values flowing
out of a basic block with more than one successor. This approach allows us to
hoist instructions to a basic block with multiple successors. The $\chi$ node in
a basic block $B$ has CHIArgs with three entries: a) the instruction (I) CHIArg
refers to, b) the basic block $Dest$ which is post-dominated by the basic block
containing I, and c) the GVN of I.

\begin{lstlisting}
struct CHIArg {
  // Edge destination (shows the direction of flow),
  // may not be where the I is.
  BasicBlock *Dest;
  // The instruction CHI-arg refers to.
  Instruction *I;
  // The global value number of I.
  GVNType VN;
};
\end{lstlisting}

A $\chi$ keeps information about `congruent' values flowing out of a basic
block.  It is similar to $\phi$ \cite{ssa} but in the inverse graph, and used
for trackig out-flowing values on each edge. An illustration of $\chi$ is shown
in Figure~\ref{fig:chi-intro}. The \GVN{} for both I1 and I2 is V, the $\chi$
node will save the instruction as well as the edge where the value is flowing
to.  Where $Arg1$, $Arg2$ are of type CHIArg. The $\chi$ for both I1 and I2 is
inserted in their PDF (A).

\begin{figure}
  \includegraphics[scale=0.55]{chi-example.pdf}
  \vspace*{-1.0cm}
\caption{I1, I2 are instructions in B, and C respectively, Arg1 = \{B, I1,
  V\} and Arg2 = \{C, I2, V\}, V is the value number for both I1, and I2}
\label{fig:chi-intro}
\end{figure}

A similar idea has been used in SSAPRE \cite{ssapre} for partial redundancy
elimination in spare form. When SSAPRE inserts $\Phi$ (expression PHIs), it does
so at places where (potentially) multiple expressions may merge. If a $\Phi$ has
a missing entry, that means the expression is partially available. Subsequent
algorithms work on $\Phi$s to find out redundancies. We extend the idea for
\GCM{} by propagating the anticipability in both directions.

\begin{figure}
  \includegraphics[scale=0.55]{chi-example2.pdf}
  \vspace*{-1.0cm}
\caption{I1, I2 are instructions in B, and C respectively, Arg1 = \{B, I1, V\}
  and Arg2 = \{C, I2, V\}, V is the value number for both I1, and I2. A has
  missing entry in $\chi$ so V is not anticipable}
\label{fig:chi-intro2}
\end{figure}

For code hoisting we insert a $\chi$, to a basic block with multiple
successors. Then for each instruction, we register its flow of value to its
corresponding $\chi$ which can be found in its PDF. Finally, we start walking
the CFG and remove the $\chi$ if it has a missing entry (see
Figure~\ref{fig:chi-intro2}), that means the expression is not anticipable in
the basic block having incomplete $\chi$. For the remaining $\chi$s
anticipabilty of `values' is ensured by construction. For code-sinking we insert
$\Phi$s in basic blocks with multiple predecessors. Then for each instruction we
register its flow of value to its correponding $\Phi$ which can be found in its
DF.  Finally, we start walking the CFG and remove the $\Phi$ if it has a missing
entry (see Figure~\ref{fig:phi-intro}), that means the expression is not
available in the basic block having incomplete $\Phi$. For the remaining $\Phi$s
availability of `values' is ensured by construction.

\begin{figure}[h]
  \includegraphics[scale=0.55]{phi-example.pdf}
  \vspace*{-1.0cm}
  \caption{I1, I2 are instructions in B, and C respectively, Arg1 = \{B, I1, V\}
  and Arg2 = \{C, I2, V\}, V is the value number for both I1, and I2. A has
  missing entry in $\Phi$ so V is not available}
\label{fig:phi-intro}
\end{figure}

In order to limit the number of spurious $\chi$s and $\Phi$s,
\cite{ssapre-improv}, we insert them only for values with multiple occurrences
in the function as they are the interesting candidates for this \GCM{}. We
believe this factorization of anticipability/availability of expressions may be
suitable for faster implementation of other global (across basic blocks)
analyses/optimizations as well. Before we consider factors affecting the
legality of \GCM{}, we introduce the concept of barriers.

\subsection{Barriers}
\label{subsec:barriers}
Barriers are based on the concept of pinned instructions \cite{click1995global}
but extended to adapt to \LLVM{} IR. A basic-block in \LLVM{} IR is actually an
extended basic-block \cite{ebb} because there might be non-returning calls in
the middle of a basic-block. There might be other instructions like stores,
where compiler may not guarantee progress of control flow, all such instructions
are marked as barriers. A barrier will prevent any instruction occurring
lexically below it from getting hoisted. On the other hand, a sink barrier will
prevent any instruction above it from getting sunk. If there are no barriers in
a basic block, any instruction satisfying other legality checks can be sunk.

\begin{lstlisting}[caption={Compute hoist and sink barriers},captionpos=b]
// Compute both hoist and sink barriers
void computeBarriers()
  for each basic-block B in a function:
    barrier_found = false
    for each instruction I in B:
      if I does not guarantee progress:
        mark I as a barrier instruction
        barrier_found = true
        break;

    // Find the last barrier below which instructions
    // can be sunk.
    if barrier_found:
      for each instruction I in B in reverse order:
        if I does not guarantee progress:
          mark I as a sink barrier
          break
\end{lstlisting}

In the absence of context, as in our current implementation, some instructions
which might be safe are still classified as barriers, e.g., calls with missing
attributes.  Computing barriers allows for efficient removal of non-movable
candidates, this makes the legality check converge faster.

\subsection{Legality of movable instructions}
\label{subsec:legality}
Since the equality of movable candidates is purely based on the value numbers,
we also need to establish if hoisting them or sinking them would be legal. The
completeness of $\chi$ and $\Phi$ are necessary for \GCM{} but not
sufficient. This is because congruence of value numbers only implies that they
compute same value if the inputs are same. So for instructions like memory
references, they may compute different values should the memory get modified
along the path. For that we need to check for intersecting side-effects along
the path.

We need to prove that on all execution paths from the initial position of an
instruction to its destination, the side effects appear in the same order. In
general, we cannot introduce a new computation along any path of execution
without checking for undefined behavior \cite{undef}. It is also necessary to
check if there are indirect branch targets, e.g., landing pad, case statements,
or other forms of goto labels along the path because it becomes difficult to
prove safety checks in those cases.

Scalars are the easiest to hoist because we only need to check for availability
of operands. For sinking the scalars we need to make sure that there are no uses
of the instruction along the path. For instructions with memory references like
loads, stores and calls, Memory SSA allows efficient way of iterating over the
use-def chains of memory references on a factored graph. Hoisting loads/stores
across calls also requires precise analysis of all the memory addresses accessed
by the call. The current implementation being an intraprocedural pass, cannot
schedule aggressively across calls. In the presence of pure calls, loads can be
hoisted but stores can't. Also, if a call throws exceptions, or if it may not
return, nothing can be scheduled across that call.

For sinking, higher ranked expressions would be sunk first. And it would
be illegal to sink higher ranked identical expressions if they are not fully
anticipable in the common post-dominator. For example:

\begin{lstlisting}
B0: i0 = load B
    if p is true then goto B1 else goto B2
B1: i1 = load A
    c1 = i1 + 10
    d1 = i0 + 20
    goto B3
B2: i2 = load A
    c2 = i2 + 10
    d2 = i0 + 20
    goto B3

B3: PHIc1, c2)
    PHI(d1, d2)
\end{lstlisting}

In this example (c1, c2) or (d1, d2) are potential sinkable candidates.  (c1,
c2) depend on i1 and i2 respectively which are also in their original basic
blocks; c1 and c2 are not fully anticipable in B3. So without knowing the
ability to sink `i1' and `i2' it would be illegal to sink (c1, c2) to B3. On the
other hand (d1, d2) can safely be sunk because their operands are readily
available at the sink point, i.e., B3. It should also be noted that, even if the
expressions are identical and operands are available, it still requires a both
instructions to be post-dominated by a unique $\Phi$ to be legally sinkable.

\subsection{Profitability check (Cost models)}
\label{subsec:cost-models}
After the legality checks have passed, we check if a \GCM{} is profitable.  That
takes into account the impact \GCM{} would have on various parameters that might
affect runtime performance, e.g., impact on live-range, gain in the code
size. The current implementation makes effort to not regress in performance and
still reduce register spills. We implement a cost model to reduce register
pressure.

\subsubsection{Reduce register pressure}
\label{hoist:reg-pressure}
Hoisting upwards will decrease the live-range of its use, if it is a last use (a
kill) but increase the live-range of its definition. Conversely, sinking will
decrease the live-range of the defined register but increase the live-range for
killed operands. If the live-range after \GCM{} will decrease, it will be
moved. Essentially, if there is one killed operand, code hoisting will either
decrease or preserve the register pressure as long as the hoisting of killed
operands does not cross a use.  Similarly, code-sinking (of an instructions
which defines a register) will either decrease or preserve the register pressure
as long as there is one operand killed at most and the instruction is not sunk
below the killing use of a non-killed operand. The following example explains
how \GCM{} can reduce the register pressure.

\begin{lstlisting}
B0:
    b = m
    c = n
    if p is true then goto B1 else goto B2
B1:
    a0 = b<kill> + c<kill>
    goto B3
B2:
    a1 = b<kill> + c<kill>
    goto B3
B3: ...
\end{lstlisting}

After hoisting, $a0$ and $a1$ are removed and a copy of $a0$ is placed in
$B0$. Since $b$ and $c$ are killed in both $a0$ and $a1$, hoisting the
expressions will reduce the register pressure because two registers will be
freed at the insertion point but only one register will be required to assign
the definition of hoisted instruction $a0$. We have reduced performance
regressions with this cost model even if LLVM has a live-range splitting
\cite{cooper1998live} optimization. Sinking may also reduce the register
pressure when the use operands are not kills.

\subsubsection{Hoisting an expression across a call}
\label{cost:across-calls}
Call expressions are difficult to analyze in an intraprocedural optimization
pass like \GCM{}. Even hoisting scalars across calls is tricky because it can
increase the number of spills. During the register allocation of call
expressions, the argument registers (also known as the caller saved registers)
are saved because they might be modified by the callee and after the call they
are restored. Due to this frame setup the register pressure is high because the
number of available registers are reduced by the number of caller saved
registers. In that situation a computation that reduces register pressure may be
beneficial to hoist. Similarly, a computation that introduces a register
(without killing any other operand) before a call may be beneficial to sink.

\section{A fast algorithm for code motion of congruent computations}
\label{subsec:optimistic}
Once all the legality and profitability checks are satisfied for a set of
congruent instructions, they are suitable candidates for code motion. A copy of
the computation is inserted at the hoisting/sinking point along with any
instructions which needed to be rematerialized. Thereafter, all the computations
made redundant by the new copy are removed and affected data structures are
updated.

\begin{lstlisting}[caption={Global Code Motion},captionpos=b]
void doGCM()
  Analyses available:
    Dominator Tree, DFS Numbering, Memory SSA
  Compute DJ-graph of function
  constructMergeSet(CFG)
  computeBarriers()
  do:
    Compute GVN of each expression in the function
    // Repeat hoisting if any candidates were hoisted.
  while (doCodeHoisting() > 0)
  // Code-sinking only once
  doCodeSinking()
\end{lstlisting}

\GCM{} basically consists of two parts, i.e., hoisting and sinking. This
implementation only moves congruent instructions which makes it easier to ensure
that we do not introduce a redundant computation along any path of execution. If
the dependency of a hoistable candidate is in the same basic-block as the
candidate, then the dependency must also be hoistable, otherwise hoisting will
be illegal or would require a complicated code generation to make it legal. The
current algorithm discards cases if the dependency is neither hoistable nor
rematerializable. We collect the \GVN{} of all the instructions in the function
and iterate on the list of instructions having identical \GVN{}s. The algorithm
\emph{doGCM} prefers hoisting (\emph{doCodeHoisting}) to sinking
(\emph{doCodeSinking}).

\begin{lstlisting}[caption={Hoist instructions},captionpos=b]
int doCodeHoisting(VNs)
  // VNs = map (VN -> list of I with value VN)
  Sort VNs in ascending order

  for each VN in VNs:
    for each instruction with the same VN:
      Find its post-dominance-frontier (PDF)
      Insert the CHI-node at PDF

  for each $\chi$-nodes in the CFG:
    for each $\chi$-arg in $\chi$-node:
      Remove $\chi$-arg if it fails legality checks

  for each $\chi$-nodes in the CFG:
    if $\chi$ does not have missing entries:
      Perform code hoisting
      update data structures and statistics

  return number of hoisted instructions
\end{lstlisting}

Code hoisting opens new opportunities for other higher ranked
\cite{rosen1988global} (depended on candidates which got hoisted) which are
encountered subsequently as the algorithm visits instructions in the increasing
order of their ranks. Some of the congruent higher ranked computations do not
get equal value numbers so we rerun the code-hoisting algorithm until there are
no more instructions left to be hoisted. The hoisting converges quickly with 2-3
iterations.

\begin{lstlisting}[caption={Sink instructions},captionpos=b]
int doCodeSinking()
  // VNs = map (VN -> list of I with value VN)
  Sort VNs in ascending order

  for each VN in VNs:
    for each instruction with the same VN:
      Find its dominance-frontier (DF)
      Insert the $\Phi$-node at DF

  for each $\Phi$-nodes in the CFG:
    for each $\Phi$-arg in PHI-node:
      Remove $\Phi$-arg if it fails legality checks

  for each $\Phi$-nodes in the CFG:
    if $\Phi$ does not have missing entries:
      Perform code sinking
      if a $\phi$ has same entries, reassign
      update data structures and statistics

  return number of sinked instructions
\end{lstlisting}

After no more candidates are hoistable, sinking is performed. It is only
performed once because it may not open new opportunities for sinkable
candidates. For sinking we iterate on the $\Phi$ nodes. After each code-motion
the \SSA{} form is restored by updating the intermediate representation (IR) to
reflect the changes. At the same time \MemorySSA{} is updated as well. Both the
data structures can be updated very efficiently. After one full iteration of
hoisting, the \GVN{} needs to be recomputed to explore new hoistable candidates.
Finally after the transformation is done, we verify a set of post-conditions to
establish that program invariants are maintained: e.g., consistency of use-defs,
and \SSA{} semantics.

\subsection{Time complexity of algorithm}
The complexity of code hoisting is linear in the number of instructions that are
candidates for \gcm{}, matching the complexity of \PRE{} on \SSA{} form.  The
analyses computed for this pass like Global Value Numbering, Marking Barriers,
are both linear in number of instructions in a function. Computation of DJ-graph
and merge-sets is $O(E)$, E being number of edges in the CFG. Liveness analysis
is not very expensive \cite{das2012} but still only performed on-demand for
hoistable candidates. Other analyses like Alias Analysis, \MemorySSA{} and
Dominator Tree are already available in the \LLVM{} pass pipeline.

Although we recompute \GVN{}, and Barriers for each iteration of the \gcm{},
there is still gain in the compilation times (see
Section~\ref{tab:compile-time}). We have also provided appropriate compiler
flags to expedite \gcm{} by bailing out with fewer iterations, or skip the
liveness based profitability analysis to aggressively move the code as long as
they are legal.

\section{Experimental Evaluation}
\label{sec:experimental-results}
% Experimental results: Register spills, compile time, run time. compile-time comparison w.r.t. gcc
For evaluation of \gcm{}, we built \SPEC{} with and without the \GCM{}
optimization. All the experiments were conducted on an \xlinux{} machine and at
-Ofast optimization level with inlining and loop unrolling disabled; inliner and
loop-unroller vary wildly with \GCM{} optimization so it was difficult to
visualize the impact. Each benchmark was run three times and the best result was
taken. We collected execution time and code-size which is listed in
Table~\ref{tab:code-size}, other compiler statistics are listed in
Table~\ref{tab:stats}, Table~\ref{tab:spills}, and
Figure~\ref{tab:code-motion-metric}.  We also measured compile time as shown in
Table~\ref{tab:compile-time}. The benchmark 464.h264ref caused an internal
compiler error with the trunk clang so it could not be reported.

\begin{table}[h!]
  \begin{center}
    \scalebox{0.7}{
    \begin{tabular}{|l|c|c|c|}
      \hline
% TODO: Check if putting code-size or spills number is okay or only %age is required.
% llvm trunk vs. llvm trunk with code-motion
                &\% performance uplift & size reduction (Bytes) \\
Benchmarks      & at -Ofast (high better)& at -Ofast (high better)\\\hline
400.perlbench	&  1.95  & 984   \\\hline
401.bzip2	& -0.84  & 16    \\\hline
403.gcc	        & -2.30  & 2128  \\\hline
429.mcf	        & -4.36  & 16    \\\hline
433.milc	& -1.86  & 48    \\\hline
444.namd	& -2.40  & 544   \\\hline
445.gobmk	& -0.52  & 56    \\\hline
447.dealII	&  0.39  & -2560 \\\hline
450.soplex	& -0.24  & -32   \\\hline
453.povray	& -0.05  & 848   \\\hline
456.hmmer	&  0.93  & 296   \\\hline
458.sjeng	& -0.58  & 232   \\\hline
462.libquantum	&  8.17  & -16   \\\hline
470.lbm	        &  2.57  & 0     \\\hline
471.omnetpp	&  5.30  & 152   \\\hline
473.astar	& -0.17  & 48    \\\hline
482.sphinx3	& -0.11  & 16    \\\hline
483.xalancbmk	& -2.98  & 392   \\\hline
    \end{tabular}}
  \end{center}
  \caption{Execution time (ratio) and code size change (Bytes) with and without
    \GCM{} on \SPEC{}}
  \label{tab:code-size}
\end{table}

Overall, we can see some good improvements in 462.libquantum, 471omnetpp, etc.
and some regressions in 429.mcf, 483.xalancbmk, etc. The \gcm{} pass was run once
in the pass pipeline (at -Ofast), after local common subexpression elimination
(EarlyCSE) pass. Because EarlyCSE removes local redundancies, \GCM{} would have
to analyze less redundant instructions. We believe some performance regressions
can be mitigated by running the \GCM{} few more times.

We see some code size reduction in 400.perlbench, 403.gcc, and 453.povray as
shown in Table~\ref{tab:code-size}.  The code size listed here is the text size
delta (from size command in linux) of the final executable for each benchmark.  On
the other hand 447.dealII increased in code size. This is because once the
code-size of a function decreases (due to \GCM{}) it becomes cheaper to do some
optimizations like inlining, loop unrolling, copy propagation, rematerialization
etc. We tried to minimize the noise by disabling inliner, and loop unroller but
still we can see some increase in the code size.

To see the impact of \GCM{} on inliner we ran \GCM{} with inliner enabled and
collected the compile-time statistics produced by LLVM (See
Table~\ref{tab:stats}) . Except for 447.dealII, 450.soplex, 482.sphinx3 and
483.xalancbmk all other benchmarks got more functions inlined. Also, since the
pass runs early, it affects many optimizations which rely on the number of
instructions, length of the use-def chain, and other metrics.

\begin{table}[h!]
  \begin{center}
    \scalebox{0.8}{
    \begin{tabular}{|l|r|r|r|r|}
      \hline
      &\multicolumn{2}{|c|}{Loop Invariant Motion}  & \multicolumn{2}{|c|}{Functions} \\\hline
 Benchmarks       & Hoisted\% & Sunk\% &  Inlined\%   & Deleted\% \\\hline
 400.perlbench	  & 2.1 & -2.8 &  0.2 &  0.9  \\\hline
 401.bzip2	  & 14.4& 0.0  &  4.7 &  3.1  \\\hline
 403.gcc	  & 13.7& 27.3 &  0.9 &  0.1  \\\hline
 429.mcf	  & -7.3& 0.0  &  0.0 &  0.0  \\\hline
 433.milc	  & 10.2& 0.0  &  8.0 &  0.0  \\\hline
 444.namd	  & 0.1 & 0.0  &  0.8 &  -3.5 \\\hline
 445.gobmk	  & 0.6 & 5.6  &  5.4 &  1.7  \\\hline
 447.dealII	  & 62.2& -86.7& -0.2 &  -0.8 \\\hline
 450.soplex	  & 3.9 & -5.6 & -0.2 &  -0.4 \\\hline
 453.povray	  & -4.8& 39.1 &  0.2 &  -0.1 \\\hline
 456.hmmer	  & 1.0 & 166.7&  0.0 &  -1.9 \\\hline
 458.sjeng	  & 11.5& 0.0  &  6.3 &  0.0  \\\hline
 462.libquantum   & 6.6 & 0.0  &  0.0 &  0.0  \\\hline
 470.lbm	  & 0.0 & 0.0  &  0.0 &  0.0  \\\hline
 471.omnetpp	  & 17.0& -6.3 &  1.1 &  -0.4 \\\hline
 473.astar	  & 9.1 & 0.0  &  1.1 &  0.0  \\\hline
 482.sphinx3	  & 5.9 & -18.2&  -2.0&  0.0  \\\hline
 483.xalancbmk	  & 14.6& -3.6 &  -1.0&  -1.2 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Change in the static compile time metrics w.r.t. \GCM{} on \SPEC{} at
    -Ofast. The \% columns show percentage increase in metric w.r.t. baseline,
    i.e., without \GCM{}. \GCM{} improves LICM and inlining in general.}
  \label{tab:stats}
\end{table}

Table~\ref{tab:stats}, which shows how \gcm{} impacted important compiler
optimizations like inlining and loop invariant code motion (LICM). LICM removes
loop invariant code out of the loop. It may hoist the code to the loop's
preheader or sink the code to loop's post-dominator
\cite{steven1997advanced}. As we can see number of instructions hoisted
increased significantly in several benchmarks, although number of instructions
sunk did not change by much.  Number of inlined functions also improved for most
benchmarks. In general when more functions were inlined, more functions were
deleted as in 401.bzip2, 445.gobmk.

\begin{table}[h!]
  \begin{center}
    \scalebox{0.7}{
    \begin{tabular}{|l|r|r|r|}
      \hline
Benchmarks        &base(-Ofast) &\GCM{}(-Ofast) &GCM/base\\\hline
400.perlbench	  & 2539	& 2493	& 0.98 \\\hline
401.bzip2	  & 718	        & 707	& 0.98 \\\hline
403.gcc	          & 5782	& 5722	& 0.99 \\\hline
429.mcf	          & 14	        & 17	& 1.21 \\\hline
433.milc	  & 635	        & 642	& 1.01 \\\hline
444.namd	  & 3223	& 3274	& 1.01 \\\hline
445.gobmk	  & 2166	& 2173	& 1.00 \\\hline
447.dealII	  & 11074	& 10234	& 0.92 \\\hline
450.soplex	  & 1122	& 1125	& 1.00 \\\hline
453.povray	  & 4701	& 4692	& 1.09 \\\hline
456.hmmer	  & 1197	& 1274	& 1.06 \\\hline
458.sjeng	  & 168	        & 176	& 1.05 \\\hline
462.libquantum	  & 118	        & 118	& 1.00 \\\hline
470.lbm	          & 33	        & 33	& 1.00 \\\hline
471.omnetpp	  & 524	        & 527	& 1.00 \\\hline
473.astar	  & 180	        & 201	& 1.12 \\\hline
482.sphinx3	  & 674	        & 670	& 0.99 \\\hline
483.xalancbmk	  & 5102	& 4965	& 0.97 \\\hline
Grand Total	  & 43349	& 42497	& 0.98 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Number of spills with and without\GCM{} on SPEC2006 at
    -Ofast. Lower is better in GCM/base. Spills reduced by 2\%.}
  \label{tab:spills}
\end{table}

Spills are listed in Table~\ref{tab:spills}. Spills were reduced by an average
of 2\% on the \SPEC{} benchmark suite. This is a result of cost-model which
limits the \gcm{} to those candidates which do not increase the register
pressure (except for loads Section~\ref{subsec:cost-models}) overall spills
still decreases.

\begin{table}[h]
  \begin{center}
    \scalebox{0.7}{
    \begin{tabular}{|l|c|c|c|}
      \hline
      Benchmarks      &   Baseline	        & \GCM{}           & \%Decrease\\
       & (in Millions)& (in Millions)& (lower is better)\\\hline
  400.perlbench	  &   186	& 184	& 0.99 \\\hline
  401.bzip2	  &   76	& 75	& 0.99 \\\hline
  403.gcc         &   1,079	& 1,073	& 0.99 \\\hline
  429.mcf         &   78	& 77	& 0.99 \\\hline
  433.milc	  &   272	& 269	& 0.99 \\\hline
  444.namd	  &   78	& 77	& 0.99 \\\hline
  445.gobmk	  &   261	& 258	& 0.99 \\\hline
  447.dealII	  &   643	& 638	& 0.99 \\\hline
  450.soplex	  &   245	& 242	& 0.99 \\\hline
  453.povray	  &   499	& 494	& 0.99 \\\hline
  456.hmmer	  &   214	& 211	& 0.99 \\\hline
  458.sjeng	  &   89	& 88	& 0.99 \\\hline
  462.libquantum  &   84	& 84	& 0.99 \\\hline
  470.lbm         &   71	& 71	& 1.00 \\\hline
  471.omnetpp	  &   382	& 379	& 0.99 \\\hline
  473.astar	  &   78	& 77	& 0.99 \\\hline
  482.sphinx3	  &   158	& 156	& 0.99 \\\hline
  483.xalancbmk	  &   22,485	& 22,455& 1.00 \\\hline
    \end{tabular}}
  \end{center}
  \caption{Change in total number of instructions executed during the
    compilation with and without \GCM{} on \SPEC{} at -Ofast. The
    instructions were counted using valgrind.}
  \label{tab:compile-time}
\end{table}

The compilation time actually reduced for most of the benchmarks as a result of
code motion.  This is because, as \gcm{} removes (congruent) instructions, the
number of instructions to be processed by later compilation passes also
reduces. The compilation time listed here is the total instruction count
executed during the compilation of each benchmark as output from `valgrind
-{}-tool=cachegrind' \cite{valgrind}. We used valgrind because it was a simple
and reliable way to show the impact of our optimization.

\begin{figure*}[h!]
  \includegraphics[width=0.8\textwidth,height=8cm]{gcm-stats.pdf}
  \vspace*{-1.5cm}
\caption{\GCM{} stats on SPEC2006 at -Ofast. Loads are hoisted the most followed
  by scalars, stores and calls.}
\label{tab:code-motion-metric}
\end{figure*}

The compile time metrics of \GCM{} are listed in
Figure~\ref{tab:code-motion-metric}. The table lists the number of scalars,
loads, stores and calls hoisted as well as removed. For each category, the
number of instructions removed is greater or equal to the number of instructions
hoisted because each code-motion is performed only when at least one identical
computation is found. Loads are hoisted the most followed by scalars, stores and
calls in decreasing order.  This was the common trend in all our
experiments. One reason why loads are hoisted the most is the early execution of
this pass (before mem2reg pass which scalarizes some memory references) in the
\LLVM{} pass pipeline. Passes like mem2reg and instcombine might actually remove
those loads and the number of hoisted loads may change should the \GCM{} pass be
scheduled later. We can see only a few sunk instructions because hoisting is
performed before sinking (Section~\ref{subsec:optimistic}) and some of the
candidates which can either be hoisted or sunk got hoisted.

\section{Conclusions and Future Work}
\label{sec:future-work}
We have presented a fast algorithm to perform \gcm{} of congruent computations
in SSA form. We introduced data structures like $\chi$ nodes, $\Phi$ nodes which
helps exploit flow of values in sparse representation.  Experimental results
show that it is a useful performance and code-size optimization. It also enables
inlining and LICM opportunities, reduces register spill. To preserve performance
and not hoist/sink too much, we have implemented a register pressure aware cost
model as described in Section~\ref{subsec:cost-models}. This pass does not hoist
fully redundant expressions because they are already removed by a later partial
redundancy elimination pass in LLVM.

We would like to explore other places in the optimization pipeline where \GCM{}
could improve performance. It may be useful to integrate \GCM{} with \LLVM{'}s
new GVN-PRE implementation such that more congruent instructions can be
scheduled. This would also expose more redundancies to the GVN-PRE. Also, since
\GCM{} runs very early in the pass pipeline, it will be good to evaluate the
code size/performance impact when it is run in sync with \GVN{}-\PRE{} just like
\GCC{} does. With the implementation of \GCM{} in \LLVM{}, the passes which rely
on the code-size or instruction-counts to make optimization decisions need to be
revisited for example, the inliner, the loop unroller etc.

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
We would like to thank Daniel Berlin for his code reviews and for his feedback
on earlier versions of this paper and Brian Grayson for motivating examples that
started this work.
\end{acks}

%% Bibliography
\bibliography{Bibliography}
%\bibliography{bibfile}

%% Appendix
%\appendix
%\section{Appendix}
%Text of appendix \ldots

\end{document}